---
title: "Multi-omics Analysis Workflow with MultiAssayExperiment"
output:
  html_document:
    df_print: paged
---
A demonstration of data management and analysis of multiple biological assays.
We'll begin by installing the necessary packages and loading their respective libraries.
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("MultiAssayExperiment")

BiocManager::install("RaggedExperiment")

BiocManager::install("curatedTCGAData")

BiocManager::install("GenomicDataCommons")

BiocManager::install("SummarizedExperiment")

BiocManager::install("SingleCellExperiment")

BiocManager::install("TCGAutils")

BiocManager::install("UpSetR")

BiocManager::install("mirbase.db")

BiocManager::install("AnnotationFilter")

BiocManager::install("EnsDb.Hsapiens.v86")

BiocManager::install("survival")

BiocManager::install("survminer")

BiocManager::install("pheatmap")
```

```{r}
library(MultiAssayExperiment)
library(GenomicRanges)
library(RaggedExperiment)
library(curatedTCGAData)
library(GenomicDataCommons)
library(SummarizedExperiment)
library(SingleCellExperiment)
library(TCGAutils)
library(UpSetR)
library(mirbase.db)
library(AnnotationFilter)
library(EnsDb.Hsapiens.v86)
library(survival)
library(survminer)
library(pheatmap)
```
The TCGA dataset is provided with the MultiAssayExperiment package. There were 92 patients. All assays were NOT performed on every patient.
Five assays included:
1. RNASeqGeneNorm: gene mRNA abundance by RNA-seq
2. gistict: GISTIC genomic copy number by gene
3. RPPAArray: protein abundance by Reverse Phase Protein Array
4. Mutations: non-silent somatic mutations by gene
5. miRNASeqGene: microRNA abundane by microRNA-seq

```{r}
#import dataset
data("miniACC")
```

```{r}
miniACC
```

Examine the data.
```{r}
colData(miniACC)[1:4, 1:4]
```
```{r}
#select features of interest and display as a table
table(miniACC$race)
```
```{r}
table(miniACC$days_to_death)
```
Explore the range of experiments included in the dataset. 
```{r}
experiments(miniACC)
```
Use sampleMap to provide graphical representation of relationship between biological units and experimental results.
The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
```{r}
sampleMap(miniACC)
```

